An Invariant Framework for Conducting Reproducible Computational Science

### 2015-01-12 evolution.tex
Framework

Given the challenges of reproducing HEP applications, we now describe components of a framework that must be present for other researchers to effectively reproduce the application. 
The framework consists of the following components (also shown in Figure \ref{}:

@@

\item {\bf Capture} In audit phase, PTU captures the execution of the application into a package. The resulting package contains the source code, its dependencies (system files, and files on network such as from CernVM-FS) and the data consumed and written by the application.

\item{\bf Preserve} The generated package is transferred to a Docker container of choice, instantiated from a vanilla, bare-bones image from a particular release (eg., CentOS 6.5, Fedora, or Ubuntu).  The container provides a base execution environment for the package. Now, the container with the package can be saved as a docker image.

\item {Distribute} The saved image can be stored and distributed in a repository like Docker Hub. We imagine these repositories would be preserved, and linked with a digital library.

 There are a number of other considerations: is it practical to store the data with the image? What if the “application” requires many processing cores to complete its task? Can one preserve the “DNA” of the application in one small, practical step, yet store everything needed for future reproductions? xHow do we minimize the complexity of the image repository (hub), stemming unnecessary growth and profitably curating the collection?

For purposes of preservation validation, a logical preservation unit (PLU) might be defined which consists of a minimal execution of the software using a small, test data sample, with specified outputs for comparison with future reproduction-validation runs.

A DASPOS preservation testing environment might consist of:

A collection of base images to be used when creating preserved platform-packages
A tool set that assists users in the 1+2+3 steps outlined above
A registry/repo service
A validation service consisting of a number of container slots that can be started on demand
A nightly testing service (Jenkins-like) that continuously monitors the reproducibility state of the preserved application + dependencies + environment + platform.

>>>>

Given the many challenges of reproducing HEP applications, we now describe a formal framework, which if present within large collaborations, can enable application developers to share their 
application with other researchers, and for other researchers to reproduce the shared application. In effect, the framework itself should be 
be invariant to changes in who owns the application (the developer or the reproducer), where it was built and compiled. To satisfy invariance, the framework must include mechanisms for:

@@

\item {\bf Capturing Dependencies and Configurations:} Capturing tools must record dependencies that are used by the program, including hardware, OS, kernels, static and dynamic dependencies, local and networked depedencies, source codes and data files. Stateful interactions with commercial software, such as proprietary databases and which cannot be captured due to licensing agreements must be persisted so that they can replayed later without the presence of the commercial software.  In effect, a captured application should behave exactly the same way as the application developer intended it to be. 
% In audit phase, PTU captures the execution of the application into a package. The resulting package contains the source code, its dependencies (system files, and files on network such as from CernVM-FS) and the data consumed and written by the application.

\item{\bf Preservation of Captured Entities:} By preservation we imply appropriate mechanisms for (i) documentation of the application development, and (b) automation of any task that becomes inevitably 
necessary for repeating the application in exactly the same way as the developer executed it.

Documentation and specification during application development can be a onerous. The preservation framework must make literate programming tools available that 
focus less on documentation, but script, integrate and execute the dependencies so that they are resolved as part of documentation. 
Automation can extend to various tasks that are necessary for ensuring repeatability such as building software, provisioning of hardware, 
validation of software against security fixes, new features, or deprecation of old software, and even monitoring the reproducibility state of a preserved application, i.e., its source code, dependencies, environment, and platform.
Automated builds and provisioning and continuous integration service can significantly lower the barrier to run the application in a new environment.

In spite of the preservation mechanisms, the application software may not run as intended. For a reproducer's understanding, it may also be useful to include 
a \emph{logical preservation unit} (PLU) that consists of a minimal execution of the software using a small, test data sample, and with specified outputs.
The provenance of this PLU must be captured so that the reproducer can make comparison with future reproduction-validation runs.

\item {\bf Distribution of Preserved Packages:} A captured and preserved application must be persistently stored and distributed through a repository. We imagine these repositories to be themselves preserved, and linked with a digital library. Metadata and flexible annotation should be part of this repository for curation over time.
